---
title: "ECO 935 Homework 1: Ahmed Almezail"
output: md_document
---


Q1


```{r, echo=FALSE}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(scales)
library(RColorBrewer)

```

```{r, echo=FALSE}
ABIA = read.csv('../Assignments/ABIA.csv')

```

First lets play with our data and breakdown the delays

```{r, echo=FALSE}
#summary of Arrival delays in general
ABIA %>%
  group_by(ArrDelay) %>%
  summarize(avg_ArrDelay = mean(ArrDelay),
            min_ArrDelay = min(ArrDelay),
            max_ArrDelay = max(ArrDelay)) %>%
  head(5) %>%
  round(1)

#summary of delays going to Austin
To_AUS_delays = ABIA %>%
  group_by(ArrDelay + DepDelay) %>%
  filter(Dest == "AUS") %>%
  summarize(avg_ArrDelay = mean(ArrDelay),
            min_ArrDelay = min(ArrDelay),
            max_ArrDelay = max(ArrDelay)) %>%
  head(5) %>%
  round(1)

#summary of delays departing from Austin
From_AUS_delays = ABIA %>%
  group_by(ArrDelay + DepDelay) %>%
  filter(Origin == "AUS") %>%
  summarize(avg_ArrDelay = mean(ArrDelay),
            min_ArrDelay = min(ArrDelay),
            max_ArrDelay = max(ArrDelay)) %>%
  head(5) %>%
  round(1)


```
  
  

Now we want to focus on those 7 variables

```{r, echo=FALSE}
D = ABIA %>%
  select(Month, DayofMonth, UniqueCarrier, Origin, Dest, ArrDelay, DepDelay)

summary(D)
```


So which airline with the most departure delays?

```{r, echo=FALSE}
D_delay = D %>%
  select(DepDelay, UniqueCarrier, Month) %>%
  filter(!is.na(DepDelay) & !is.na(UniqueCarrier)) %>%
  group_by(UniqueCarrier) %>%
  summarise(n = length(UniqueCarrier), total = sum(DepDelay), avrdelay = mean(DepDelay)) %>%
  arrange(desc(n)) %>%
  data.frame() %>%
  top_n(20, wt=n)


ggplot(D_delay, aes(x=reorder(UniqueCarrier, n), y=n, label=round(avrdelay, digits = 0))) + 
  geom_bar(stat="identity", fill='red') +
  coord_flip() +
  geom_text(position=position_dodge(width=0.5), hjust=-0.25) +
  labs(title = "Total Departure Delays by Airline",
       x = "Airline",
       y = "Total Delay",
       subtitle = "Average Delay Value shown to the right of bar",
       caption = "Southwest and American Airlines are the highest airlines with total departure delays")

```



No wonder American Airlines is the 2nd on the list.

Now let's jump to the delay of Arrival flights.

```{r, echo=FALSE}
A_delay = D %>%
  select(ArrDelay, UniqueCarrier, Month) %>%
  filter(!is.na(ArrDelay) & !is.na(UniqueCarrier)) %>%
  group_by(UniqueCarrier) %>%
  summarise(n = length(UniqueCarrier), total = sum(ArrDelay), avrdelay2 = mean(ArrDelay)) %>%
  arrange(desc(n)) %>%
  data.frame() %>%
  top_n(20, wt=n)

ggplot(A_delay, aes(x=reorder(UniqueCarrier, n), y=n, label=round(avrdelay2, digits = 0))) + 
  geom_bar(stat="identity", fill='red') +
  coord_flip() +
  geom_text(position=position_dodge(width=0.5), hjust=-0.25) +
  labs(title = "Total Arrival Delays by Airline",
       x = "Airline",
       y = "Total Delay",
       subtitle = "Average Delay Value shown to the right of bar",
       caption = "Southwest and American Airlines are the highest airlines with total arrival delays")

```





Departure delays per airline by months:

```{r, echo=FALSE}
D_delay100 = D %>% 
  filter(DepDelay > 100) 

ggplot(D_delay100) +
  geom_point(mapping = aes(x= Month, y = DepDelay, color = UniqueCarrier)) +
  xlim(0,12)+
  facet_wrap(~UniqueCarrier, nrow=4) +
  labs(title = "Total Departure Delays by Airline by Month",
       x = "Month",
       y = "Total Delay",
       colour = "Airline Name",
       caption = "Most of the delays per airline is in the middle of the year.
            This graph should give you a hint in which airline to choose based on the
       date of your flight from Austin.")

```



Arrival delays per airline by months:

```{r, echo=FALSE}
A_delay100 = D %>% 
  filter(DepDelay > 100) 

ggplot(A_delay100) +
  geom_point(mapping = aes(x= Month, y = ArrDelay, color = UniqueCarrier)) +
  xlim(0,12)+
  facet_wrap(~UniqueCarrier, nrow=4) +
  labs(title = "Total Arrival Delays by Airline by Month",
       x = "Month",
       y = "Total Delay",
       colour = "Airline Name",
       caption = "Most of the delays per airline is in the middle of the year.
            This graph should give you a hint in which airline to choose based on the
       date of your flight to Austin.")
```





Now let's analyze the top 2 airlines with highest total delays:

Southwest
Departure:

```{r, echo=FALSE}
D_WN = D %>% 
  filter(UniqueCarrier=="WN" & !is.na(Origin)& !is.na(Origin)) %>%   
  group_by(Origin) %>%
  summarise(n = length(Origin), total = sum(DepDelay), avrdelay3 = mean(DepDelay)) %>%
  arrange(desc(n)) %>%
  data.frame() %>%
  top_n(20, wt=n)

ggplot(D_WN, aes(x=reorder(Origin, n), y=n, label=round(avrdelay3, digits = 0))) + 
  geom_bar(stat="identity", fill="blue") +
  coord_flip() +
  geom_text(position=position_dodge(width=0.5), hjust=-0.25) + 
  labs(title = "Total Departure Delays by Origin Airport for Southwest Airlines",
       x = "Origin Airport",
       y = "Total Delay",
       subtitle = "Average Delay Value shown to the right of bar",
       caption = "Since the data is about flights from and to Austin, it is expected to have Austin on the top.
       However most other delays are going to Dallas, Houstn, and Pheonix.")

```



Southwest Airlines
Arrival:

```{r, echo=FALSE}
A_WN = D %>% 
  filter(UniqueCarrier=="WN" & !is.na(Origin)& !is.na(Origin)) %>%   
  group_by(Origin) %>%
  summarise(n = length(Origin), total = sum(ArrDelay), avrdelay5 = mean(ArrDelay)) %>%
  arrange(desc(n)) %>%
  data.frame() %>%
  top_n(20, wt=n)

ggplot(A_WN, aes(x=reorder(Origin, n), y=n, label=round(avrdelay5, digits = 0))) + 
  geom_bar(stat="identity", fill="blue") +
  coord_flip() +
  geom_text(position=position_dodge(width=0.5), hjust=-0.25) + 
  labs(title = "Total Departure Delays by Origin Airport for Southwest Airlines",
       x = "Origin Airport",
       y = "Total Delay",
       subtitle = "Average Delay Value shown to the right of bar",
       caption = "Since the data is about flights from and to Austin, it is expected to have Austin on the top.
       However most other delays are also from Dallas, Houstn, and Pheonix.")


```


Now Lets move to American Airlines
Departure:

```{r, echo=FALSE}
D_AA = D %>% 
  filter(UniqueCarrier=="AA" & !is.na(Origin)& !is.na(Origin)) %>%   
  group_by(Origin) %>%
  summarise(n = length(Origin), total = sum(DepDelay), avrdelay3 = mean(DepDelay)) %>%
  arrange(desc(n)) %>%
  data.frame() %>%
  top_n(20, wt=n)

ggplot(D_AA, aes(x=reorder(Origin, n), y=n, label=round(avrdelay3, digits = 0))) + 
  geom_bar(stat="identity", fill="blue") +
  coord_flip() +
  geom_text(position=position_dodge(width=0.5), hjust=-0.25) + 
  labs(title = "Total Departure Delays by Origin Airport for American Airlines",
       x = "Origin Airport",
       y = "Total Delay",
       subtitle = "Average Delay Value shown to the right of bar",
       caption = "Since the data is about flights from and to Austin, it is expected to have Austin on the top.
       However most other delays are going to Dallas, Chicago, and Los Angles")

```


American Airlines
Arrival:

```{r}
A_AA = D %>% 
  filter(UniqueCarrier=="AA" & !is.na(Origin)& !is.na(Origin)) %>%   
  group_by(Origin) %>%
  summarise(n = length(Origin), total = sum(ArrDelay), avrdelay4 = mean(ArrDelay)) %>%
  arrange(desc(n)) %>%
  data.frame() %>%
  top_n(20, wt=n)

ggplot(A_AA, aes(x=reorder(Origin, n), y=n, label=round(avrdelay4, digits = 0))) + 
  geom_bar(stat="identity", fill="blue") +
  coord_flip() +
  geom_text(position=position_dodge(width=0.5), hjust=-0.25) + 
  labs(title = "Total Departure Delays by Origin Airport for American Airlines",
       x = "Origin Airport",
       y = "Total Delay",
       subtitle = "Average Delay Value shown to the right of bar",
       caption = "Since the data is about flights from and to Austin, it is expected to have Austin on the top.
       However most other delays are going to Dallas and Chicago.")
```





I hope now you have a better idea in how to plan for your spring break. Make sure you don't miss any minute.



Q2



```{r, echo=FALSE}
library(tidyverse)
library(ggplot2)
library(data.table)
library(dplyr)
```


```{r, echo=FALSE}
billboard = read.csv('../Assignments/billboard.csv')

```


PartA


```{r, echo=FALSE}
billboard = billboard %>%
  select(performer, song, year, week, week_position, song_id, weeks_on_chart)

top_songs = billboard %>%
  group_by(performer, song)
yes = top_songs %>%
    tally() %>%
    arrange(desc(n)) %>%
    data.frame() %>%
    top_n(10, wt=n)
  
  colnames(yes)[3] <- "count"
  
  view(yes)
```

  
  
  As we can see, Radioactive by Imagine dragons is the most popular song according to the billboard.
  
  PartB
  
  
```{r, echo=FALSE}
div_songs = billboard %>%
    group_by(performer, song, year) %>%
  filter_all(all_vars(.!=1958 & .!=2021))
  view(div_songs)
  
  group_by_year <- 
    data.table(div_songs)[ , .(unique_songs = length(unique(song))), 
                     by = year] 
  
  ggplot(group_by_year, aes(x=year)) + 
    geom_line(aes(y = unique_songs), color = "darkred") +
    labs(x="Year",
         y="number of unique songs",
         title="Musical Diversity",
         subtitle = "The number of unique songs appeared on the Billboard by the year",
         caption = "We can notice how in the 60's, the music diversity was increasing
         reaching 800 uniqe songs in a year. However, the number started to fall sharply
         from the late 60's to the end of the century with almost half of the peak value giving it sharp negative trend. Afterward,
         with the boom of the Pop music in the US, the diversity surged again reaching almost 600 songs in 2010.
         Then it fall again by about 100 in magnitude by 2014, then bounced back reaching the 800 level by 2020.
         This might be because of the rise of the social media, and the strong music marketing industry recently.")
```
  

  
  
  
  PartC
 
```{r, echo=FALSE}
hit_songs = billboard %>%
   select(weeks_on_chart, performer, song) %>%
   group_by(performer, song) %>%
   summarize(count = n()) %>%
   arrange(desc(count)) %>%
   filter(count>=10)

 view(hit_songs)
 
 singers = hit_songs %>%
   group_by(performer) %>%
   summarize(count = n()) %>%
   filter(count >= 30) %>%
   arrange(desc(count))

 view(singers) 

 ggplot(singers) +
   geom_col(aes(fct_reorder(performer, count), count)) +
   coord_flip(expand = TRUE) +
   labs(x="Singers",
        title = "10-Week Hit!",
      caption = "As we can see, there are 19 singers who were able to have at least
      30 songs for a 10-week hit on the Billboard. Feel sorry for Imagine Dragons not to be in this list.")
  
  



```
 
 
 Q3
 
```{r, echo=FALSE}
 library(tidyverse)
library(ggplot2)
library(dplyr)
library(matrixStats)

 olympics_top20 = read.csv('../Assignments/olympics_top20.csv')

```
 
 


1st part

```{r, echo=FALSE}
height_percentile = olympics_top20 %>%
  select(sex, height, sport) %>%
  group_by(height) %>%
  filter(sex == "F" & sport == "Athletics") %>%
  summarise(quantile(olympics_top20$height, 0.95))

head(height_percentile)

Female_95_Percentile = 197

Female_95_Percentile

```
So The 95th percentile of height for females in the Atheleics games is 197


  
2nd part

```{r, echo=FALSE}
t =  olympics_top20 %>%
  select(sex, height, event) %>%
  group_by(event) %>%
  filter(sex == "F" & str_detect(event, 'Single')) %>%
  data.frame() 


##count no. of events after filter
t1 = xtabs(~event, data=t) %>%
  data.frame()

#200 m Sd

t200 = t %>%
  filter(event == "Canoeing Women's Kayak Singles, 200 metres") 


w200mSd = sd(t200$height)

#500 m Sd

t500 = t %>%
  filter(event == "Canoeing Women's Kayak Singles, 500 metres") 


w500mSd = sd(t500$height)

# Slalom

tslalom = t %>%
  filter(event == "Canoeing Women's Kayak Singles, Slalom") 


wslalomSd = sd(tslalom$height)


# Rowing

trowing = t %>%
  filter(event == "Rowing Women's Single Sculls") 


wrowingSd = sd(trowing$height)

Stdev = c(w200mSd, w500mSd, wslalomSd, wrowingSd)


Stdev
```

So, Canoeing Women's Kayak Singles 500 metres event has the highest standard deviation in height


part 3

```{r, echo=FALSE}
zf =  olympics_top20 %>%
  group_by(year, age) %>% 
  filter(sport == "Swimming" & sex == "F") %>%
  data.frame() 

setDT(zf)

zm =  olympics_top20 %>%
  select(sex, age, sport, year) %>%
  group_by(year, age) %>% 
  filter(sport == "Swimming" & sex == "M") %>%
  data.frame() 

setDT(zm)

zfmean = zf[ ,list(mean=mean(age)), by=year]

zmmean = zm[ ,list(mean=mean(age)), by=year]




setDF(zfmean)
setDF(zmmean)

colnames(zfmean)[2] <- "Female_Avg_Age"
colnames(zmmean)[2] <- "Male_Avg_Age"

Age_Mean = zmmean %>% full_join(zfmean)


ggplot(Age_Mean, aes(x=year)) + 
  geom_line(aes(y = Female_Avg_Age), color = "darkred") +
  geom_line(aes(y = Male_Avg_Age), color="steelblue") +
  labs(x="Year",
       y="Avg Age",
       title="Avg Age of Swimmers Across the Years",
         subtitle = "Red is Female Blue is Male",
       caption = "We can notice that, The sport was only for males before it started to be for females as well regulrly in the 50's.
       However, it was played once for females in 1924 then purse the game eventually in the mid of the century.
       The chart excluded the observations in 1924 to not do any predictions for the gapped era.
       The average age of male has increased drastically from 19 to 32 in jus 25 years, followed by a sharp
       decline in the next few years. Then the average start to increase gradually along with females 
       reaching about 24 years old for men and about 23 years old for women.")


 
```

As it shown in the graph, we can notice that, The sport was only for males before it started to be for females as well regularly in the 50's.However, it was played once for females in 1924 then purse the game eventually in the mid of the century.

The chart excluded the observations in 1924 to not do any predictions for the gapped era.
 The average age of male has increased drastically from 19 to 32 in jus 25 years, followed by a sharp decline in the next few years. Then the average start to increase gradually along with females reaching  about 24 years old for men and about 23 years old for women.
 
 
 Q4
 
```{r, echo=FALSE}
library(tidyverse)
library(ggplot2)
library(rsample)  # for creating train/test splits
library(caret)
library(modelr)
library(parallel)
library(foreach)
library(dplyr)

sclass = read.csv('../Assignments/sclass.csv')


```
 
Filter the data by trim
 
```{r, echo=FALSE}


class350 = sclass %>%
  select(price, trim, mileage) %>%
  filter(trim == "350")

class63AMG = sclass %>%
  select(price, trim, mileage) %>%
  filter(trim == "63 AMG")

summary(class350)
summary(class63AMG)
```






plot the data for both trims

```{r, echo=FALSE}
ggplot(data = class350) + 
  geom_point(mapping = aes(x = mileage, y = price), color='darkgrey') +
 labs(title = "Price of 350 S Class by Milages")

ggplot(data = class63AMG) + 
  geom_point(mapping = aes(x = mileage, y = price), color='darkgrey') +
  labs(title = "Price of 63 AMG S Class by Milages")

```




Make a train-test split for both trims

```{r, echo=FALSE}
class350_split =  initial_split(class350, prop=0.8)
class350_train = training(class350_split)
class350_test = testing(class350_split)

class63AMG_split =  initial_split(class63AMG, prop=0.8)
class63AMG_train = training(class63AMG_split)
class63AMG_test  = testing(class63AMG_split)
```



Fit a linear model for both trims

```{r, echo=FALSE}
lm1_350 = lm(price ~ mileage, data=class350_train)
coef(lm1_350)

ggplot(data = lm1_350) + 
  geom_point(mapping = aes(x = mileage, y = price)) +
  geom_abline(intercept = 71541.86, slope = -0.5804595, color='red') +
  labs(title = "Fitted model for 350 S Class")

lm1_63AMG = lm(price ~ mileage, data=class63AMG_train)
coef(lm1_63AMG)

ggplot(data = lm1_63AMG) + 
  geom_point(mapping = aes(x = mileage, y = price)) +
  geom_abline(intercept = 71541.86, slope = -0.5804595, color='red') +
  labs(title = "Fitted model for 63AMG S Class")

```








Now lets try KNN model with K = 2, 5, 10, 20, 30, 35, 40, 45, 50, 55, 60, 80, 100, 150, 200 

First for 350 S Class and we will view the RMSE

```{r}
knn2_350 = knnreg(price ~ mileage, data=class350_train, k=2)
rmse(knn2_350, class350_test)

knn5_350 = knnreg(price ~ mileage, data=class350_train, k=5)
rmse(knn5_350, class350_test)

knn10_350 = knnreg(price ~ mileage, data=class350_train, k=10)
rmse(knn10_350, class350_test)

knn20_350 = knnreg(price ~ mileage, data=class350_train, k=20)
rmse(knn20_350, class350_test)

knn30_350 = knnreg(price ~ mileage, data=class350_train, k=30)
rmse(knn30_350, class350_test)

knn35_350 = knnreg(price ~ mileage, data=class350_train, k=35)
rmse(knn35_350, class350_test)

knn40_350 = knnreg(price ~ mileage, data=class350_train, k=40)
rmse(knn40_350, class350_test)

knn45_350 = knnreg(price ~ mileage, data=class350_train, k=45)
rmse(knn45_350, class350_test)

knn50_350 = knnreg(price ~ mileage, data=class350_train, k=50)
rmse(knn50_350, class350_test)

knn55_350 = knnreg(price ~ mileage, data=class350_train, k=55)
rmse(knn55_350, class350_test)

knn60_350 = knnreg(price ~ mileage, data=class350_train, k=60)
rmse(knn60_350, class350_test)

knn80_350 = knnreg(price ~ mileage, data=class350_train, k=80)
rmse(knn80_350, class350_test)

knn100_350 = knnreg(price ~ mileage, data=class350_train, k=100)
rmse(knn100_350, class350_test)

knn150_350 = knnreg(price ~ mileage, data=class350_train, k=150)
rmse(knn150_350, class350_test)

knn200_350 = knnreg(price ~ mileage, data=class350_train, k=200)
rmse(knn200_350, class350_test)


```



Now the same for 63AMG S Class

```{r, echo=FALSE}
knn2_63AMG = knnreg(price ~ mileage, data=class63AMG_train, k=2)
rmse(knn2_63AMG, class63AMG_test)

knn5_63AMG = knnreg(price ~ mileage, data=class63AMG_train, k=5)
rmse(knn5_63AMG, class63AMG_test)

knn10_63AMG = knnreg(price ~ mileage, data=class63AMG_train, k=10)
rmse(knn10_63AMG, class63AMG_test)

knn20_63AMG = knnreg(price ~ mileage, data=class63AMG_train, k=20)
rmse(knn20_63AMG, class63AMG_test)

knn30_63AMG = knnreg(price ~ mileage, data=class63AMG_train, k=30)
rmse(knn30_63AMG, class63AMG_test)

knn35_63AMG = knnreg(price ~ mileage, data=class63AMG_train, k=35)
rmse(knn35_63AMG, class63AMG_test)

knn40_63AMG = knnreg(price ~ mileage, data=class63AMG_train, k=40)
rmse(knn40_63AMG, class63AMG_test)

knn45_63AMG = knnreg(price ~ mileage, data=class63AMG_train, k=45)
rmse(knn45_63AMG, class63AMG_test)

knn50_63AMG = knnreg(price ~ mileage, data=class63AMG_train, k=50)
rmse(knn50_63AMG, class63AMG_test)

knn55_63AMG = knnreg(price ~ mileage, data=class63AMG_train, k=55)
rmse(knn55_63AMG, class63AMG_test)

knn60_63AMG = knnreg(price ~ mileage, data=class63AMG_train, k=60)
rmse(knn60_63AMG, class63AMG_test)

knn80_63AMG = knnreg(price ~ mileage, data=class63AMG_train, k=80)
rmse(knn80_63AMG, class63AMG_test)

knn100_63AMG = knnreg(price ~ mileage, data=class63AMG_train, k=100)
rmse(knn100_63AMG, class63AMG_test)

knn150_63AMG = knnreg(price ~ mileage, data=class63AMG_train, k=150)
rmse(knn150_63AMG, class63AMG_test)

knn200_63AMG = knnreg(price ~ mileage, data=class63AMG_train, k=200)
rmse(knn200_63AMG, class63AMG_test)

```




Lets plot the fit

350 S Class

```{r, echo=FALSE}
class350_testK2 = class350_test %>%
  mutate(class350_predK2 = predict(knn2_350, class350_test))

K2_test = ggplot(data = class350_testK2) + 
  geom_point(mapping = aes(x = mileage, y = price), alpha=0.3) +
   labs(title = "Plot Graph of 350 S Class ")
 

K2_test

# K2 predictions
K2_test + geom_line(aes(x = mileage, y = class350_predK2), color='red', size=1.5) +
   labs(title = "KNN fitting model of 350 S Class for K2 ")

#K5

class350_testK5 = class350_test %>%
  mutate(class350_predK5 = predict(knn5_350, class350_test))

K5_test = ggplot(data = class350_testK5) + 
  geom_point(mapping = aes(x = mileage, y = price), alpha=0.3) 



# K5 predictions
K5_test + geom_line(aes(x = mileage, y = class350_predK5), color='red', size=1.5) +
  labs(title = "KNN fitting model of 350 S Class for K5 ")

#K10

class350_testK10 = class350_test %>%
  mutate(class350_predK10 = predict(knn10_350, class350_test))

K10_test = ggplot(data = class350_testK10) + 
  geom_point(mapping = aes(x = mileage, y = price), alpha=0.3) 



# K10 predictions
K10_test + geom_line(aes(x = mileage, y = class350_predK10), color='red', size=1.5) +
  labs(title = "KNN fitting model of 350 S Class for K10 ")

#K20

class350_testK20 = class350_test %>%
  mutate(class350_predK20 = predict(knn20_350, class350_test))

K20_test = ggplot(data = class350_testK20) + 
  geom_point(mapping = aes(x = mileage, y = price), alpha=0.3) 



# K20 predictions
K20_test + geom_line(aes(x = mileage, y = class350_predK20), color='red', size=1.5) +
  labs(title = "KNN fitting model of 350 S Class for K20 ")


#K30

class350_testK30 = class350_test %>%
  mutate(class350_predK30 = predict(knn30_350, class350_test))

K30_test = ggplot(data = class350_testK30) + 
  geom_point(mapping = aes(x = mileage, y = price), alpha=0.3) 



# K30 predictions
K30_test + geom_line(aes(x = mileage, y = class350_predK30), color='red', size=1.5) +
  labs(title = "KNN fitting model of 350 S Class for K30 ")


#K35

class350_testK35 = class350_test %>%
  mutate(class350_predK35 = predict(knn35_350, class350_test))

K35_test = ggplot(data = class350_testK35) + 
  geom_point(mapping = aes(x = mileage, y = price), alpha=0.3) 



# K35 predictions
K35_test + geom_line(aes(x = mileage, y = class350_predK35), color='red', size=1.5) +
  labs(title = "KNN fitting model of 350 S Class for K35 ")


#K40

class350_testK40 = class350_test %>%
  mutate(class350_predK40 = predict(knn40_350, class350_test))

K40_test = ggplot(data = class350_testK40) + 
  geom_point(mapping = aes(x = mileage, y = price), alpha=0.3) 



# K40 predictions
K40_test + geom_line(aes(x = mileage, y = class350_predK40), color='red', size=1.5) +
  labs(title = "KNN fitting model of 350 S Class for K40 ")


#K45

class350_testK45 = class350_test %>%
  mutate(class350_predK45 = predict(knn45_350, class350_test))

K45_test = ggplot(data = class350_testK45) + 
  geom_point(mapping = aes(x = mileage, y = price), alpha=0.3) 



# K45 predictions
K45_test + geom_line(aes(x = mileage, y = class350_predK45), color='red', size=1.5) +
  labs(title = "KNN fitting model of 350 S Class for K45 ")


#K50

class350_testK50 = class350_test %>%
  mutate(class350_predK50 = predict(knn50_350, class350_test))

K50_test = ggplot(data = class350_testK50) + 
  geom_point(mapping = aes(x = mileage, y = price), alpha=0.3) 



# K50 predictions
K50_test + geom_line(aes(x = mileage, y = class350_predK50), color='red', size=1.5) +
  labs(title = "KNN fitting model of 350 S Class for K50 ")


#K55

class350_testK55 = class350_test %>%
  mutate(class350_predK55 = predict(knn55_350, class350_test))

K55_test = ggplot(data = class350_testK55) + 
  geom_point(mapping = aes(x = mileage, y = price), alpha=0.3) 



# K55 predictions
K55_test + geom_line(aes(x = mileage, y = class350_predK55), color='red', size=1.5) +
  labs(title = "KNN fitting model of 350 S Class for K55 ")


#K60

class350_testK60 = class350_test %>%
  mutate(class350_predK60 = predict(knn60_350, class350_test))

K60_test = ggplot(data = class350_testK60) + 
  geom_point(mapping = aes(x = mileage, y = price), alpha=0.3) 



# K60 predictions
K60_test + geom_line(aes(x = mileage, y = class350_predK60), color='red', size=1.5) +
  labs(title = "KNN fitting model of 350 S Class for K60 ")

#K80

class350_testK80 = class350_test %>%
  mutate(class350_predK80 = predict(knn80_350, class350_test))

K80_test = ggplot(data = class350_testK80) + 
  geom_point(mapping = aes(x = mileage, y = price), alpha=0.3) 



# K80 predictions
K80_test + geom_line(aes(x = mileage, y = class350_predK80), color='red', size=1.5) +
  labs(title = "KNN fitting model of 350 S Class for K80 ")


#K100

class350_testK100 = class350_test %>%
  mutate(class350_predK100 = predict(knn100_350, class350_test))

K100_test = ggplot(data = class350_testK100) + 
  geom_point(mapping = aes(x = mileage, y = price), alpha=0.3) 



# K100 predictions
K100_test + geom_line(aes(x = mileage, y = class350_predK100), color='red', size=1.5) +
  labs(title = "KNN fitting model of 350 S Class for K100 ")


#K150

class350_testK150 = class350_test %>%
  mutate(class350_predK150 = predict(knn150_350, class350_test))

K150_test = ggplot(data = class350_testK150) + 
  geom_point(mapping = aes(x = mileage, y = price), alpha=0.3) 



# K150 predictions
K150_test + geom_line(aes(x = mileage, y = class350_predK150), color='red', size=1.5) +
  labs(title = "KNN fitting model of 350 S Class for K150 ")


#K200

class350_testK200 = class350_test %>%
  mutate(class350_predK200 = predict(knn200_350, class350_test))

K200_test = ggplot(data = class350_testK200) + 
  geom_point(mapping = aes(x = mileage, y = price), alpha=0.3) 



# K200 predictions
K200_test + geom_line(aes(x = mileage, y = class350_predK200), color='red', size=1.5) +
  labs(title = "KNN fitting model of 350 S Class for K200 ")


```



Now lets have a look at the fitting of 63AMG S Class model

```{r, echo=FALSE}
#K2

class63AMG_test = class63AMG_test %>%
  mutate(class63AMG_predK2 = predict(knn2_63AMG, class63AMG_test))

KK2_test = ggplot(data = class63AMG_test) + 
  geom_point(mapping = aes(x = mileage, y = price), alpha=0.3) +
  labs(title = "Plot Graph of 63AMG S Class")

KK2_test

# K2 predictions
KK2_test + geom_line(aes(x = mileage, y = class63AMG_predK2), color='red', size=1.5) +
  labs(title = "KNN fitting model of 63AMG S Class for K2")

#K5

class63AMG_test = class63AMG_test %>%
  mutate(class63AMG_predK5 = predict(knn5_63AMG, class63AMG_test))

KK5_test = ggplot(data = class63AMG_test) + 
  geom_point(mapping = aes(x = mileage, y = price), alpha=0.3) 



# K5 predictions
KK5_test + geom_line(aes(x = mileage, y = class63AMG_predK5), color='red', size=1.5) +
  labs(title = "KNN fitting model of 63AMG S Class for K5")

#K10

class63AMG_test = class63AMG_test %>%
  mutate(class63AMG_predK10 = predict(knn10_63AMG, class63AMG_test))

KK10_test = ggplot(data = class63AMG_test) + 
  geom_point(mapping = aes(x = mileage, y = price), alpha=0.3) 



# K10 predictions
KK10_test + geom_line(aes(x = mileage, y = class63AMG_predK10), color='red', size=1.5) +
  labs(title = "KNN fitting model of 63AMG S Class for K10")

#K20

class63AMG_test = class63AMG_test %>%
  mutate(class63AMG_predK20 = predict(knn20_63AMG, class63AMG_test))

KK20_test = ggplot(data = class63AMG_test) + 
  geom_point(mapping = aes(x = mileage, y = price), alpha=0.3) +
  labs(title = "KNN fitting model of 63AMG S Class for K20")



# K20 predictions
KK20_test + geom_line(aes(x = mileage, y = class63AMG_predK20), color='red', size=1.5)


#K30

class63AMG_test = class63AMG_test %>%
  mutate(class63AMG_predK30 = predict(knn30_63AMG, class63AMG_test))

KK30_test = ggplot(data = class63AMG_test) + 
  geom_point(mapping = aes(x = mileage, y = price), alpha=0.3) 



# K30 predictions
KK30_test + geom_line(aes(x = mileage, y = class63AMG_predK30), color='red', size=1.5) +
  labs(title = "KNN fitting model of 63AMG S Class for K30")



#K35

class63AMG_test = class63AMG_test %>%
  mutate(class63AMG_predK35 = predict(knn35_63AMG, class63AMG_test))

KK35_test = ggplot(data = class63AMG_test) + 
  geom_point(mapping = aes(x = mileage, y = price), alpha=0.3) 



# K35 predictions
KK35_test + geom_line(aes(x = mileage, y = class63AMG_predK35), color='red', size=1.5) +
  labs(title = "KNN fitting model of 63AMG S Class for K35")


#K40

class63AMG_test = class63AMG_test %>%
  mutate(class63AMG_predK40 = predict(knn40_63AMG, class63AMG_test))

KK40_test = ggplot(data = class63AMG_test) + 
  geom_point(mapping = aes(x = mileage, y = price), alpha=0.3) 



# K40 predictions
KK40_test + geom_line(aes(x = mileage, y = class63AMG_predK40), color='red', size=1.5) +
  labs(title = "KNN fitting model of 63AMG S Class for K40")


#K45

class63AMG_test = class63AMG_test %>%
  mutate(class63AMG_predK45 = predict(knn45_63AMG, class63AMG_test))

KK45_test = ggplot(data = class63AMG_test) + 
  geom_point(mapping = aes(x = mileage, y = price), alpha=0.3) 



# K45 predictions
KK45_test + geom_line(aes(x = mileage, y = class63AMG_predK45), color='red', size=1.5) +
  labs(title = "KNN fitting model of 63AMG S Class for K45")


#K50

class63AMG_test = class63AMG_test %>%
  mutate(class63AMG_predK50 = predict(knn50_63AMG, class63AMG_test))

KK50_test = ggplot(data = class63AMG_test) + 
  geom_point(mapping = aes(x = mileage, y = price), alpha=0.3) 



# K50 predictions
KK50_test + geom_line(aes(x = mileage, y = class63AMG_predK50), color='red', size=1.5) +
  labs(title = "KNN fitting model of 63AMG S Class for K50")


#K55

class63AMG_test = class63AMG_test %>%
  mutate(class63AMG_predK55 = predict(knn55_63AMG, class63AMG_test))

KK55_test = ggplot(data = class63AMG_test) + 
  geom_point(mapping = aes(x = mileage, y = price), alpha=0.3) 



# K55 predictions
KK55_test + geom_line(aes(x = mileage, y = class63AMG_predK55), color='red', size=1.5) +
  labs(title = "KNN fitting model of 63AMG S Class for K55")


#K60

class63AMG_test = class63AMG_test %>%
  mutate(class63AMG_predK60 = predict(knn60_63AMG, class63AMG_test))

KK60_test = ggplot(data = class63AMG_test) + 
  geom_point(mapping = aes(x = mileage, y = price), alpha=0.3) 



# K60 predictions
KK60_test + geom_line(aes(x = mileage, y = class63AMG_predK60), color='red', size=1.5) +
  labs(title = "KNN fitting model of 63AMG S Class for K60")

#K80

class63AMG_test = class63AMG_test %>%
  mutate(class63AMG_predK80 = predict(knn80_63AMG, class63AMG_test))

KK80_test = ggplot(data = class63AMG_test) + 
  geom_point(mapping = aes(x = mileage, y = price), alpha=0.3) 



# K80 predictions
KK80_test + geom_line(aes(x = mileage, y = class63AMG_predK80), color='red', size=1.5) +
  labs(title = "KNN fitting model of 63AMG S Class for K80")


#K100

class63AMG_test = class63AMG_test %>%
  mutate(class63AMG_predK100 = predict(knn100_63AMG, class63AMG_test))

KK100_test = ggplot(data = class63AMG_test) + 
  geom_point(mapping = aes(x = mileage, y = price), alpha=0.3) 



# K100 predictions
KK100_test + geom_line(aes(x = mileage, y = class63AMG_predK100), color='red', size=1.5) +
  labs(title = "KNN fitting model of 63AMG S Class for K100")


#K150

class63AMG_test = class63AMG_test %>%
  mutate(class63AMG_predK150 = predict(knn150_63AMG, class63AMG_test))

KK150_test = ggplot(data = class63AMG_test) + 
  geom_point(mapping = aes(x = mileage, y = price), alpha=0.3) 



# K150 predictions
KK150_test + geom_line(aes(x = mileage, y = class63AMG_predK150), color='red', size=1.5) +
  labs(title = "KNN fitting model of 63AMG S Class for K150")


#K200

class63AMG_test = class63AMG_test %>%
  mutate(class63AMG_predK200 = predict(knn200_63AMG, class63AMG_test))

KK200_test = ggplot(data = class63AMG_test) + 
  geom_point(mapping = aes(x = mileage, y = price), alpha=0.3) 



# K200 predictions
KK200_test + geom_line(aes(x = mileage, y = class63AMG_predK200), color='red', size=1.5) +
  labs(title = "KNN fitting model of 63AMG S Class for K200")

```


## Class63AMG


Now lets do the cross validation for both trims

```{r,echo=FALSE}

# K-fold cross validation class350


K_folds = 5

class350_folds = crossv_kfold(class350, k=K_folds)


# map the model-fitting function over the training sets
models_350 = map(class350_folds$train, ~ knnreg(price ~ mileage, k=100, data = ., use.all=FALSE))


# map the RMSE calculation over the trained models and test sets simultaneously
errs_350 = map2_dbl(models_350, class350_folds$test, modelr::rmse)


mean(errs_350)
sd(errs_350)/sqrt(K_folds)   # approximate standard error of CV error


# so now we can do this across a range of k
k_grid = c(2, 3, 4, 5, 6, 7, 8, 9, 10, 12, 15, 18, 20, 25, 30, 35, 40, 45, 50, 55, 60, 80,
           100, 150, 200)


cv_grid = foreach(k = k_grid, .combine='rbind') %dopar% {
  models_350 = map(class350_folds$train, ~ knnreg(price ~ mileage, k=k, data = ., use.all=FALSE))
  errs_350 = map2_dbl(models_350, class350_folds$test, modelr::rmse)
  c(k=k, errs_350 = mean(errs_350), std_err_350 = sd(errs_350)/sqrt(K_folds))
} %>% as.data.frame


# plot means and std errors versus k
ggplot(cv_grid) + 
  geom_point(aes(x=k, y=errs_350)) + 
  geom_errorbar(aes(x=k, ymin = errs_350-std_err_350, ymax = errs_350+std_err_350)) + 
  scale_x_log10() +
  labs(title = "RMSE per K for 350 Class",
       y = "RMSE")


### 63AMG Class


K_folds = 5

class63AMG_folds = crossv_kfold(class63AMG, k=K_folds)


# map the model-fitting function over the training sets
models_63AMG = map(class63AMG_folds$train, ~ knnreg(price ~ mileage, k=100, data = ., use.all=FALSE))


# map the RMSE calculation over the trained models and test sets simultaneously
errs_63AMG = map2_dbl(models_63AMG, class63AMG_folds$test, modelr::rmse)


mean(errs_63AMG)
sd(errs_63AMG)/sqrt(K_folds)   # approximate standard error of CV error


# so now we can do this across a range of k
k_grid = c(2, 3, 4, 5, 6, 7, 8, 9, 10, 12, 15, 18, 20, 25, 30, 35, 40, 45, 50, 55, 60, 80,
           100, 150, 200)


cv_grid_63AMG = foreach(k = k_grid, .combine='rbind') %dopar% {
  models_63AMG = map(class63AMG_folds$train, ~ knnreg(price ~ mileage, k=k, data = ., use.all=FALSE))
  errs_63AMG = map2_dbl(models_63AMG, class63AMG_folds$test, modelr::rmse)
  c(k=k, errs_63AMG = mean(errs_63AMG), std_err_63AMG = sd(errs_63AMG)/sqrt(K_folds))
} %>% as.data.frame


# plot means and std errors versus k
ggplot(cv_grid_63AMG) + 
  geom_point(aes(x=k, y=errs_63AMG)) + 
  geom_errorbar(aes(x=k, ymin = errs_63AMG-std_err_63AMG, ymax = errs_63AMG+std_err_63AMG)) + 
  scale_x_log10() +
  labs(title = "RMSE per K for 63AMG Class",
       y = "RMSE")




```




Based on the RMSE numbers, I would use K35 for 350 S Class and K80 or K100 for 63AMG Class.

```{r, echo=FALSE}
# 350 Class K20 predictions
K20_test + geom_line(aes(x = mileage, y = class350_predK20), color='red', size=1.5) +
  labs(title = "Optimal KNN Fitting Model for 350 S Class")


# 63AMG Class K100 predictions
KK100_test + geom_line(aes(x = mileage, y = class63AMG_predK100), color='red', size=1.5) +
    labs(title = "Optimal KNN Fitting Model for 63 AMG S Class")


```




So 63 AMG class has the higher optimal value of K since it has more clustered obs in different areas of the graph. Also the plot points are more scattered than for 350 Class. So, the predicted line might have more bias but less variance. 

